### 2025-01-29 scripts/Paper-analysis.Rmd
leading vs. lagging palindrome scores

>>>>

### Palindromes in leading vs. lagging

```{r}
scores_df = read.csv('../test_6mer_leading_lagging_scores.csv', header=T)
scores_average <- scores_df %>%
  rowwise() %>%
  mutate(score_avg = mean(c_across(starts_with("score")))) %>%
  ungroup() %>%
  select(kmer, score_avg)  
# Order by average score
scores_average$kmer = ordered(scores_average$kmer,
                              levels=scores_average$kmer[order(scores_average$score_avg)])
scores.plot = ggplot(scores_average, aes(kmer, score_avg))+geom_point()+
  coord_flip()

# Same for counts
counts_df = read.csv('../test_6mer_leading_lagging_counts.csv', header=T)
counts_average <- counts_df %>%
  rowwise() %>%
  mutate(count_avg = mean(c_across(starts_with("count")))) %>%
  ungroup() %>%
  select(kmer, count_avg)

# Merge together
merge_scores_counts = merge(counts_average, scores_average, by="kmer")
# average score and average count are well-correlated
ggplot(merge_scores_counts, aes(count_avg, score_avg))+geom_point()

# Order by average score
merge_scores_counts$kmer = ordered(merge_scores_counts$kmer,
                              levels=scores_average$kmer[order(scores_average$score_avg)])
scores.plot = ggplot(merge_scores_counts, aes(score_avg, kmer))+
    geom_vline(xintercept = 0)+
  geom_bar(stat="identity")+
  geom_vline(xintercept = median(merge_scores_counts$score_avg), linetype='dashed', colour='red')+
  theme_bw()+
  theme(axis.text=element_text(colour="black", size=12),
        axis.title.x=element_text(size=24),
        axis.text.x=element_text(size=18))+
  theme(axis.text.y=element_text(hjust=0))+
  ylab("")+
  xlab("Mean depletion score\n(leading vs. lagging, 10kb)")
wilcox.test(merge_scores_counts$score_avg)
ggsave(scores.plot, file='../results/6mer-palindromes-scores.pdf',
       height=10, width=10)
```
